---
title: "Session_5_Tutorial"
output: html_document
---
## Instal & Load Required Pacakges
### CRAN packages
Install CRAN packages

```{r}
cran.packages <- c("tidyverse",
                   "ggrepel",
                   "kableExtra",
                   "ggupset")

cran.load <- function(pkg){
        new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
        if (length(new.pkg)){
          install.packages(new.pkg, dependencies = TRUE)
          }
        sapply(pkg, require, character.only = TRUE)
}
cran.load(cran.packages)
```

### Bioconductor Packages
Install Biocondustor packages

```{r}
bioconductor.packages <- c("org.Hs.eg.db",
                           "clusterProfiler",
                           "enrichplot",
                           "pathview")

bioconductor.load <- function(pkg){
        new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
        if (length(new.pkg)){
          if (!requireNamespace("BiocManager", quietly = TRUE))
            install.packages("BiocManager")
            BiocManager::install(new.pkg)
        }
        sapply(pkg, require, character.only = TRUE)
}
bioconductor.load(bioconductor.packages)
```

## Load Save Data
Load the pairwise results file from the previous session, session 4, the data should be in the ../Session_4/DESeq2/Treatment_vs_Control_results/ directory and called "Treatment_vs_Control.genes.tsv" or the shrunken data would be called "lfc.Treatment_vs_Control.genes.tsv". Your working directory for today should be Session_5 so you could use the relative path "../Session_4/DESeq2/Treatment_vs_Control_results/Treatment_vs_Control.genes.tsv". The "relative" path to the file is relative to the current working directory. Use the variable name "de_results". Complete the code chunk below.
If you don't have the results file then you can get a copy from the data directory in this Session.

```{r read in previous days results file}
de_results <- 
```

Now repeat the reading in of the results file with the other file either Treatment_vs_Control.genes.tsv or lfc.Treatment_vs_Control.genes.tsv 

```{r read in previous days results file}
lfc_de_results <- 
```

## Results Directory
Create a results directory for the ClusterProfiler results files, I would call it "ClusterProfiler", but you can call it whatever you would like. 
HINT: You created a similar directory in yesterdays session. Complete the code chunk below 
```{r}
results_directory <- 
```

## Filtering
Filter the results table for genes that have a p adjusted value of 0.1 and a log fold change of 1.
NOTE: The use of the absolute value for the log2 fold change. Complete the code chunk below

```{r}
filtered_de_results <- de_results %>%
  filter(padj < qvalueCutoff,
         abs(log2FoldChange) > foldchange) %>% # Filter on p adjusted and log fold change values
  drop_na() # Remove any NA values
```

Now filter the lfc results. Complete the code chunk below

```{r}
lfc_filtered_de_results <- 
```


## Gene Ontology Analysis
Over representation Analysis
List of entrez ids for the differentially expressed genes, de_results. Use the pull() command from dplyr. Complete the code chunk below

```{r}
de_genes <- filtered_de_results %>%
```


Over representation analysis
```{r}
go_ora <- enrichGO(gene = de_genes,
                   OrgDb = org.Hs.eg.db,
                   ont = "CC", # Can choose "BP","CC", "MF" or "ALL"
                   pAdjustMethod = "BH", # Can choose "holm", "hochberg", "hommel", "bonferroni", "BH", "BY", "fdr", "none"
                   pvalueCutoff = 0.01,
                   qvalueCutoff = 0.05,
                   readable = TRUE)
```



List of entrez ids for the log fold shrunken differentially expressed genes, lfc_filtered_de_results. Use the pull() command from dplyr. Complete the code chunk below

```{r}
lfc_de_genes <- filtered_de_results %>%
```


Repeat over representation analysis with the lfc results. 
```{r}
go_ora_lfc <- 
```

Have a look at the results for the GO analyses to se if they differ.

## Gene Set enrichment Analysis

Input is a named vector of log fold changes with entrez ids as the names, sorted
```{r}
gsea_gene_list <- filtered_de_results$log2FoldChange # Get the log fold values in a character vector
names(gsea_gene_list) <- filtered_de_results$entrezgene_id # Add the entrez gene ids as the names for the log fold changes
gsea_gene_list <- sort(gsea_gene_list, decreasing = TRUE) # Sort the character vector bym logo fold change, for lowest to highest†
```

Gene set enrichment analysis

```{r}
#| echo: true
#| eval: true
go_gsea <- gseGO(gene = , # Add the gene list here
                 OrgDb = org.Hs.eg.db,
                 ont = "CC", # Can choose BP,CC, MF or ALL
                 pAdjustMethod = "BH", # Can choose "holm", "hochberg", "hommel", "bonferroni", "BH", "BY", "fdr", "none"
                 minGSSize    = 100,
                 maxGSSize    = 500,
                 pvalueCutoff = 0.01)
```

Now try using the lfc_filtered_de_results

```{r}
gsea_lfc_gene_list <-
```

Repeat geneset enrichment analysis with the lfc results.

```{r}
go_gsea_lfc <-
```


Now save the Gene Ontology results to a file. You will need to convert the results into the format of a data frame, as.data.frame() command. Use the write_delim() command and save to the results directory created earlier.

```{r}

```


Now using the dotplot() command visualise the GO analysis results, You can limit the number of results using the parameter showCategory=10. Also try visualising using the barplot() command. Or you could also visualise the lfc results.

```{r}

```

These visulaisations are ggplot2 objects and can be saved using ggsave like the other plots we have created.

## Pathway Analysis
Over representation analysis

```{r}
kegg_ora <- enrichKEGG(gene = de_genes,
                       organism = 'hsa',
                       pvalueCutoff = 0.01)
```

Save the pathway analysis to file

```{r}
kegg_ora %>%
  as.data.frame() %>%
  write_delim(file = file.path(results_directory, "ora_pathway.tsv"),
              delim = "\t")
```

Now try using the lfc_de_genes list

```{r}
kegg_ora_lfc <- 
```

Save the pathway analysis to file

```{r}

```

## Pathway Analysis
Gene set enrichment analysis

```{r}
kegg_gsea <- gseKEGG(geneList = gsea_gene_list,
                     organism = 'hsa',
                     minGSSize = 10,
                     pvalueCutoff = 0.05)
```

Save the pathway analysis to file

```{r}

```


Now try using the gsea_lfc_gene_list 

```{r}

```

## Pathway Analysis Visualisation

Create a function to retrieve the KEG pathways

```{r}
# Pathway visulaisation function
kegg_pathview <- function(kegg_pathway_id){ 
  pathview(gene.data = gsea_gene_list, # Named list of fold changes
        pathway.id = kegg_pathway_id, # KEGG pathway id
        species = "hsa", # KEG species id
        gene.idtype = "KEGG",
        kegg.native = TRUE) 
}

# List of KEGG pathway ids
kegg_gsea_list <- kegg_gsea %>%
  as.data.frame() %>%
  pull(ID) 

# Iterate over the list of pathway ids and apply the data to the pathway
lapply(kegg_gsea_list, kegg_pathview)
```
