---
title: "Session 2 - Intro to base R"
output: html_notebook
---

This is a beginners tutorial for R. It is a powerful tool for all manner of calculations, data manipulation, scientific computations and can also be used to visualise your data.

This tutorial uses R notebooks which consists of notes that are on a white back ground, "code chunks" that start with three backticks (\```) then curly brackets around the coding language, in this case R (like this {r}) and the code block end is defined by a further set of three backticks (\```). 
You can create code chunks manually or by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

The code blocks can be run by clicking on the green start arrow or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. The results of the code block will be shown below the code block.

# Base R
All the following are available in the version of R without installing any additional packages

## First the basics of R.
Basic calculations
```{r}
4 * 5 + 2
```

Try a simple calculation for your self
```{r}

```

For readability it is best to add spaces between the operators (e.g. 4 * 5), however, these are not required and can be omitted.

## Variables
Variables are an integral part of any programming language and R offers a great deal of flexibility. R does not require variable types to be declared and can take on any available data type. It can also hold any R object such as a function, the result of an analysis or a plot.
There are a number of ways to assign a value to a variable, and again, this does not depend on the type of value being assigned. 
The valid assignment operators are <- and =, with the first being preferred. 

For example, let’s save 2 to the variable x and 5 to the variable y. The arrow operator can also point in the other direction.
```{r}
x <- 2
y <- 5
x
y
```
Now we can use the variables later in our code, in this case a simple addition.
```{r}
x + y
```

Variables can hold any type of data
Characters
Note the use of the paste command, which takes the variables to be pasted together followed by the character used to separate the variables,ß in this case a white space.
```{r}
greeting <- "Hello"
name <- "Graham"

paste(greeting,name, sep = " ")
```

Vectors must contain elements of the same type for examples logical, integer, double, complex, character and raw.
This may look like a list but in R a list is a data structure that can contain elements of different types.
```{r}
vector <- c(1,2,3,4)
vector
```

The length of the vector can be determined using the length function
```{r}
length(vector)
```

Some vectors my have NA values. NA values are not equivalent to zero and any calculation with an NA value will result in NA
```{r}
z <- c(1, 2, NA, 8, 3, NA, 3)
z
sum(z) 
```

Many functions can ignore the NAs using the na.rm = TRUE feature
```{r}
sum(z, na.rm=TRUE)
```

Can add values to an existing vector using the append function
```{r}
append(z, 5) # add to the end of a vector
append(5, z) # add to the beginning of a vector
append(z,5, after = 2) # add after the second element in the vector
```

Can find out if the vector contains unique values. The unique function removes duplicated values, keeping the first occurrence of a duplicated value
```{r}
z
unique(z)
```

Can use the grepl function to find elements in a vector that match a pattern, this results in a list of booleans (TRUE or FALSE)
```{r}
gene_names <- c("ACT1","NOTCH","DMPK1")
gene_names
grepl("1",gene_names)
```

Can use that boolean list to subset the original vector
```{r}
gene_names[grepl("1",gene_names)]
```

Try this for your self, this time searching for names the contain "C".
```{r}

```

## Read in data from file

It is very common that you will need to read in data to R from a file. There are a few ways to do this, base R has a read.table() function. That allows you to read in table data with various delimiters, for example spaces, tabs or commas. Spaces are denoted like this " ". tabs by "\t" and commas by ",".
The file will be read in as a data frame. A data frame is a table or a two-dimensional array-like structure, similar to an excel spreadshet, in which each column contains values of one variable and each row contains one set of values from each column.

The code below reads in a space separated file, that has named columns,(hence the header = TRUE parameter)

```{r}
data_table <- read.table("data/example_table.txt", header = TRUE, sep = " ") # This will read in a space separated data table
data_table
```
Now try reading in th comma separated file in data/example-table.csv.
```{r}

```

The data_table was read in as a data frame, which is like a spreadsheet in excel. Data frames consist of named rows and columns, these by default are numbers. Each column can be treated like a vector and is accessed using the data frame name and the column name, separated by a dollar sign, $. For example, we can calculate the natural log for the values Sample1 in the table we just loaded.
```{r}
log(data_table$Sample1)
```

Often we would want to perform a calculation on all values in the data frame
```{r}
log(data_table)
```

If you want to perform a calculation on the whole data frame all the values have to be of the same type, e.g.all numbers (either integer or double). The error above indicates that the Gene column does not contain numeric values. We can replace the numbered row names the the Gene name column from the data frame then delete it.
```{r}
rownames(data_table) <- data_table$Gene # replace the row names with the gene names
data_table$Gene <- NULL # delete the Gene column
data_table
```
Now the data frame contains only numeric values the log function can be applied to the entire data frame
```{r}
log_table <- log(data_table)
log_table
```

Writing data to a tab separated file using the write.table() function. We have set the column names to TRUE, quote to FASLE and the separator, sep, to "\t", which is the symbol for a tab. The quote parameter is set to FALSE,  because write.table() by default surrounds the values in the data frame with double quotes, which can be annoying when using the file later.
```{r}
write.table(log_table,"data/example_log_table.tsv", col.names = TRUE, quote = FALSE, sep = "\t")
```

Try for your self write the data to a comma separated file called "example_log_table.csv"
```{r}

```


## Functions
In R you create functions which are blocks of code that can take parameters. Functions can perform very simple tasks but normally you would write a function to perform more complex computations.
A simple example of what a function looks like, below
```{r}
my_function <- function(x){
  result <- (x + 5) * 3
  return(result)
}

my_function(5)
```
Notice that when you run the function for the first time it appears in the Functions section in the Environment pane. So you can call to function later in your code without have to re-write the code or re-run the function. Functions are useful when you will be running the same computation multiple times in different parts of your code, they make the code easier to read (especially if your functions have descriptive names) and maintain.
```{r}
my_function(3)
```

Functions are also useful if you want to do the same computation on more than one element at a time. We use the R command lapply() 
```{r}
num_list <- c(4,6,8,9,2)
function_results <- lapply(num_list, my_function)
function_results
```
Notice that the function_results in the Environment pane is in the Data section along with data frames. That is because Lists are a more complex structure than vectors.
List can be converted to a vector using the unlist() function.
```{r}
unlist(function_results)
```

Write you own simple function and test it.
```{r}

```

## Getting help
R and RStudio has a built in help pages for the standard R packages and the packages that you install. These help pages will have a short description of the command/packages function with descriptions of the parameters and usage examples.
To access these help pages you can either click on the "Help" tab in the files pane on the right and enter the command you need help with into the search box in the top right of the pane. Another way to get the help pages is to type a question mark followed by the command name you need help with, e.g. ?sum. This can be in either the Console pane on in a code chunk, see below.
```{r}
?sum
```

Try this for another command you would like to know more about, for example "summary"

