---
title: "Session 2 - Intro to R"
output: html_notebook
---

This is a beginners tutorial for R. It is a powerful tool for all manner of calculations, data manipulation, scientific computations and can also be used to visualise your data.

This tutorial uses R notebooks which consists of notes that are on a white back ground, "code chunks" that start with three backticks (`) then curly brackets around the coding language, in this case R (like this {r}) and the code block end is defined by a further set of three backticks(`). 
You can create code chunks manually or by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

The code blocks can be run by clicking on the green start arrow or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. The results of the code block will be shown below the code block.

## First the basics of R.
Basic calculations
```{r}
4 * 5 + 2
```

For readability it is best to add spaces between the operators (e.g. 4 * 5), however, these are not required and can be omitted.

## Variables
Variables are an integral part of any programming language and R offers a great deal of flexibility. R does not require variable types to be declared and can take on any available data type. It can also hold any R object such as a function, the result of an analysis or a plot.
There are a number of ways to assign a value to a variable, and again, this does not depend on the type of value being assigned. 
The valid assignment operators are <- and =, with the first being preferred. 

For example, let’s save 2 to the variable x and 5 to the variable y. The arrow operator can also point in the other direction.
```{r}
x <- 2
y <- 5
x
y
```
Now we can use the variables
```{r}
x + y
```
Variables can hold any type of data
Characters
Note the use of the paste command. This is my preferred way to join variables or text.
```{r}
greeting <- "Hello"
name <- "Graham"
print(paste(greeting,name, sep = " "))

```
Lists
```{r}
list <- c(1,2,3,4)
list
```

```{r}
z <- c(1, 2, NA, 8, 3, NA, 3)
z %>% mean(na.rm=TRUE)
```

## Read in data from file

It is very common that you will need to read in data to R from a file. There are a few ways to do this, base R has a read.table() function. That allows you to read in table data with various delimiters, for example spaces, tabs or commas. Spaces are denoted like this " ". tabs by "\t" and commas by ",".
The file will be read in as a data frame. A data frame is a table or a two-dimensional array-like structure, similar to an excel spreadshet, in which each column contains values of one variable and each row contains one set of values from each column.

The code below reads in a space separated file, that has named columns,(hence the header = TRUE parameter)

```{r}
table <- read.table("data/example_table.txt", header = TRUE, sep = " ") # This will read in a space separated data table
table
```
Now try reading in th comma separated file in data/example-table.csv.
```{r}

```


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

